Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities in depth

### 2017-11-06 manuscript/manuscript.tex

\begin{summary}
We present a new open-soruce software that allows to compute the gravitational fields generated by any tesseroid whose density varies arbitrarily in depth, i.e. a forward gravity model in spherical coordinates for varying densities on the radial direction.
It is written in Python and Cython language and is compatible with the classes and functions of the library for geophysical modelling Fatiando a Terra.
It numerically approximates the gravitational fields by applying the Gauss-Legendre Quadrature along with a preexisting adaptive discretization algorithm and a new density-based discretization algorithm that control both the accuracy and the computation time of the forward model.
The first one is controlled by a predefined distance-size ratio ($D$) and reduces the errors due to the distance between the tesseroid and the computation point.
While the second one, regulated by a delta ratio ($\delta$), decreases the errors introduced by the variation of the density function.
We have also obtained analytical solutions for a spherical shell with variable density and compared them with the results of the numerical model in case of a linear, an exponential and a discontinuous density function.
These comparisons allowed to obtain default values for the distance-size and the delta ratio in order to guarantee the accuracy of the forward model.
The resulting optimal values of $D$ for the gravity potential, the components of its gradient and the Marussi tensor components are 1, 2 and 8, respectively.
While a $\delta=0.2$ is needed even for high varying exponential densities and discontinuous functions.
We have finally tested the new software by modelling the Neuqu\'en sedimentary basin, located to the east of the Andes between 32$^\circ$S and 40$^\circ$S latitude, assigning a linear density variation to it and computing all the gravity fields it generates.
\end{summary}

@@

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{figures/neuquen-basin.pdf}
\caption{(a) Topography of Neuqu\'en Basin (in km),
         (b) Thickness of the sedimentary basin (in meters),
         (c)-(h) Computed gravity fields: potential $V$ (in J/kg), gradient components $g_x$, $g_y$ and $g_z$ (in mGal) and Marussi tensor components $g_{xy}$ and $g_{zz}$ (in Eötvös), calculated at 50km of height over the ellipsoid.}
\label{fig:neuquen-basin}
\end{figure*}

@@

We must also define a density function for the entire model.
\citet{Sigismondi2012} proposed a minimum and a maximum density for the Neuqu\'en basin of 275kg/m$^3$ and 412kg/m$^3$, respectively.
So we have chosen, for this specific example, a linear density that assumes the minimum value on the surface and the maximum at 5858m, i.e. the bottom of the basin.

@@

Finally, we have shown a real data example: we have computed the gravity fields generated by the Neuqu\'en sedimentary basin through a tesseroid model with linear increasing density with depth on a regular grid at 50km height.

>>>>

\begin{summary}
We present a new open-soruce software that computes the gravitational fields generated by any tesseroid whose density varies arbitrarily in depth, i.e. a forward gravity model in spherical coordinates for varying densities on the radial direction.
It is written in Python and Cython language and is compatible with the classes and functions of the library for geophysical modelling Fatiando a Terra.
It numerically approximates the gravitational fields by applying the Gauss-Legendre Quadrature along with a preexisting adaptive discretization algorithm and a new density-based discretization algorithm that control both the accuracy and the computation time of the forward model.
The first algorithm is regulated by a predefined scalar called distance-size ratio ($D$) and reduces the errors due to the distance between the tesseroid and the computation point.
While the second one, controlled by a delta ratio ($\delta$), decreases the errors introduced by the variation of the density function.
We have also obtained analytical solutions for a spherical shell with variable density and compared them with the results of the numerical model in case of a linear, an exponential and a discontinuous density function.
These comparisons allowed to obtain default values for the distance-size and the delta ratio in order to guarantee the accuracy of the forward model.
The resulting optimal values of $D$ for the gravity potential, the components of its gradient and the Marussi tensor components are 1, 2 and 8, respectively.
While a $\delta=0.2$ is needed in the computation of every gravity field, guaranteeing the accuracy even for high varying exponential densities and discontinuous functions.
We have finally tested the new software by modelling the Neuqu\'en sedimentary basin, located to the east of the Andes between 32$^\circ$S and 40$^\circ$S latitude, assigning an exponential density variation to it and computing all the gravity fields it generates.
\end{summary}

@@

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{figures/neuquen-basin.pdf}
\caption{
         Forward model example with real data: Nequ\'en sedimentary modelled through tesseroids with exponential density.
         (a) Topography of Neuqu\'en Basin (in km),
         (b) Thickness of the sedimentary basin (in meters),
         (c)-(h) Computed gravity fields: potential $V$ (in J/kg), gradient components $g_x$, $g_y$ and $g_z$ (in mGal) and Marussi tensor components $g_{xy}$ and $g_{zz}$ (in Eötvös), calculated at 50km of height over the ellipsoid.}
\label{fig:neuquen-basin}
\end{figure*}

@@

We must also define a density function for the entire model.
\citet{Sigismondi2012} proposed a minimum and maximum density contrast for the Neuqu\'en basin of -412kg/m$^3$ and -275kg/m$^3$, respectively. \todo{ver que esos valores son de contraste y tienen q ir negativos. poner los valores absolutos.}
So we have chosen, for this specific example, an exponential density (like the one in equation \ref{eq:density-exp}) that assumes the minimum value on the surface and the maximum at 5858m, i.e. the bottom of the basin, with a value of $b$ equal to 3km.

@@

Finally, we have shown a real data example: we have computed the gravity fields generated by the Neuqu\'en sedimentary basin through a tesseroid model with exponential density with depth on a regular grid at 50km height.
